Handling String-Number Conversion in String Constraint Solving

### 2019-11-20 evaluation.tex
Update evaluation.tex
update evaluation and minor update to intro

\begin{table*}[t]
\centering
\caption{Results of z3-Trau, cvc4, and z3 on string benchmarks (numbers with * will be updated later)}
\begin{tabular}{l r | r r r r r r r}
\hline
\multicolumn{2}{c}{}                   & {\tool} & CVC4  &    Z3 & Z3-str3 & Trau+ & ABC & Ostrich \\ \hline
\multirow{3}{*}{PyEx}		& sat      & 19586*   & 19763* & 18359 &   3024*  & 19149 & 0 & 111 \\ 
							& unsat    &  3858*   &  3834* &  3851 &   3839*  & 3828 & 0 & 871 \\
							& $\times$ &  1977*   &  1824* &  3211 &  18558*  & 2444 & 25421 & 24439 \\ \hline
\multirow{3}{*}{APLAS}		& sat      &   128*   &   205* &    19 &     38*  & 132 & 289 & 0 \\
							& unsat    &   287*   &   211* &   100 &    111*  & 82 & 2 & 1 \\
							& $\times$ &   185*   &   174* &   481 &    451*  & 386 & 309 & 599 \\ \hline
\multirow{3}{*}{LeetCode}	& sat      &   856*   &   860* &   881 &    670*  & 778 & 0 & 158 \\
							& unsat    &  1784*   &  1785* &  1785 &   1780*  & 1827 & 0 & 1618 \\
							& $\times$ &    26*   &    21* &     0 &    216*  & 61 & 2666 & 890 \\ \hline
\multirow{3}{*}{StringFuzz}	& sat      &   502*   &   677* &   311 &    493*  & 688 & 439 & 0 \\
							& unsat    &   294*   &   240* &   205 &    190*  & 339 & 158 & 0 \\
							& $\times$ &   269*   &   148* &   549 &    382*  & 38 & 468 & 1065 \\\hline
\multirow{3}{*}{cvc4\textsubscript{pred}} & sat & 16* & 11* &   12 &      8*  & 247 & 316 & 21 \\
							& unsat    &   814*   &   818* &   808 &    772*  & 500 & 443 & 17 \\
							& $\times$ &     5*   &     6* &    15 &     55*  & 88 & 76 & 797 \\ \hline
\multirow{3}{*}{cvc4\textsubscript{term}} & sat & 13* & 8* &     6 &     16*  & 311 & 576 & 45 \\
							& unsat    &  1030*   &   936* &  1022 &    958*  & 596 & 349 & 22 \\
							& $\times$ &     2*   &   101* &    17 &     71*  & 137 & 120 & 978 \\ \hline
\multirow{3}{*}{SLOG}  		& sat      & 0* & 1309* & 0 & 0* & 1228 & 1036 & 1299 \\
							& unsat    & 3391* & 2082* & 0 & 2205* & 2079 & 1972 & 2079 \\
							& $\times$ & 0* & 0* & 3391 & 1186* & 84 & 383 & 13 \\ \hline
\end{tabular}
\label{table:base_benchmark}
\end{table*}

\begin{table}[t]
\centering
\caption{Results of {\tool}, cvc4, and z3 on str\_int benchmark (numbers with * will be updated later)}
\scalebox{0.7}{
\begin{tabular}{l r | r r r r}
\hline
\multicolumn{2}{c}{}                   & {\tool} & CVC4   &    Z3  & Z3-str3 \\ \hline
\multirow{3}{*}{StrInt}		& sat      &   3294*  &  2185* &  2731* &    422* \\ 
							& unsat    &  17088*  & 16377* & 16832* &  16034* \\
							& $\times$ &   1191*  &  3011* &  2010* &   5117* \\ \hline \hline
\multirow{3}{*}{leetcode-addStrings}	& sat & 622*  &  330* &  100* &  85* \\ 
							& unsat    &  1054*  & 780* & 1043* &  634* \\
							& $\times$ &  2*  &  568* &  535* & 959* \\ \hline
\multirow{3}{*}{leetcode-add\_Binary}	& sat & 520*  &  530* &  531* &  11* \\ 
							& unsat    &  1351*  & 1091* & 1091* &  1094* \\
							& $\times$ &  11*  &  261* &  260* & 777* \\ \hline
\multirow{3}{*}{leetcode-numDecodings}	& sat & 166*  &  107* &  151* &  50* \\ 
							& unsat    &  273*  & 269* & 283* &  269* \\
							& $\times$ &  92*  &  155* &  97* & 212* \\ \hline
\multirow{3}{*}{leetcode-restoreIpAddresses}	& sat & 811*  &  462* &  864* &  55* \\ 
							& unsat    &  13648*  & 13540* & 13648* &  13342* \\
							& $\times$ &  54*  &  511* &  1* & 1116* \\ \hline
\multirow{3}{*}{leetcode-validIPAddress}	& sat & 54*  &  57* &  58* &  8* \\ 
							& unsat    &  19*  & 19* & 19* &  19* \\
							& $\times$ &  4*  &  1* &  0* & 50* \\ \hline
\multirow{3}{*}{leetcode-validWordAbbreviation}	& sat & 183*  &  193* &  188* &  8* \\ 
							& unsat    &  24*  & 11* & 21* &  16* \\
							& $\times$ &  39*  &  42* &  37* & 222* \\ \hline
\multirow{3}{*}{lib-datetime\_parse\_hh\_mm\_ss\_ff}	& sat &  133*  & 133* &  133* &  88* \\ 
							& unsat    &  41*  & 41* & 41* &  43* \\
							& $\times$ &  0*  &  0* &  0* & 43* \\ \hline
\multirow{3}{*}{lib-datetime\_parse\_isoformat\_date}	& sat & 32*  &  32* &  32* &  23* \\ 
							& unsat    &  0*  & 0* & 0* &  0* \\
							& $\times$ &  0*  &  0* &  0* & 9* \\ \hline
\multirow{3}{*}{lib-distutils\_get\_build\_version}	& sat & 19*  &  19* &  19* &  4* \\ 
							& unsat    &  24*  & 24* & 24* &  24* \\
							& $\times$ &  0*  &  0* &  0* & 15* \\ \hline
\multirow{3}{*}{lib-email\_parsedate\_tz}	& sat & 68*  &  72* &  72* &  30* \\ 
							& unsat    &  138*  & 138* & 138* &  138* \\
							& $\times$ &  4*  &  0* &  0* & 42* \\ \hline
\multirow{3}{*}{lib-http\_parse\_request}	& sat & 24*  &  24* &  24* &  8* \\ 
							& unsat    &  9*  & 9* & 9* &  7* \\
							& $\times$ &  0*  &  0* &  0* & 18* \\ \hline
\multirow{3}{*}{lib-ipaddress\_ip\_int\_from\_string}	& sat & 595*  &  204* &  480* &  18* \\ 
							& unsat    &  427*  & 373* & 431* &  357* \\
							& $\times$ &  1006*  &  1451* &  1117* & 1653* \\ \hline
\multirow{3}{*}{lib-nntplib\_parse\_datetime}	& sat & 4*  &  8* &  0* &  0* \\ 
							& unsat    &  0*  & 0* & 0* &  0* \\
							& $\times$ &  4*  &  0* &  8* & 8* \\ \hline
\multirow{3}{*}{lib-smtpd\_parseargs}	& sat & 27*  &  27* &  27* &  24* \\ 
							& unsat    &  72*  & 72* & 72* &  66* \\
							& $\times$ &  0*  &  0* &  0* & 9* \\ \hline
\multirow{3}{*}{lib-wsgiref\_check\_status}	& sat & 10*  &  10* &  10* &  6* \\ 
							& unsat    &  9*  & 9* & 9* &  9* \\
							& $\times$ &  0*  &  0* &  0* & 4* \\ \hline
% \multirow{4}{*}{filtered\_str\_int}	& sat      &   3281*  &  2210* &  2729* &    424* \\
% 									& unsat    &   2912*  &  2211* &  2655* &   1944* \\
% 									& $\times$ &   1203*  &  2975* &  2012* &   5028* \\ \hline
\end{tabular}}
\label{table:str_int_benchmark}
\end{table}

@@

Our tool {\tool} is done by implementing the proposed string constraint solving procedure on top of the SMT solver Z3~\cite{z3} as a theory solver. In this way we can concentrate on solving constraints connected by conjunctions and let Z3 to handle other boolean connectives. It also enable us to solve not only formulas over string constraints, but also the combination of string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulas.

@@

We perform two sets of experiments. In the first experiment, we compare with other tools on benchmarks over basic string constraints. We use the following benchmark examples in this experiments:
\begin{itemize}
	\item PyEx~\cite{pyex} is from running the symbolic executor PyEx over some Python packages.
	\item APLAS~\cite{aplas} includes 600 hand-drafted tests consisting of only equality and integer length constraints.
	\item LeetCode is from running PyEx over some sample code collected from the LeetCode~\cite{LeetCode} website.
	\item StringFuzz~\cite{StringFuzz}  is generated from the fuzz testing tool of the same name.
	\item $\text{cvc4}_{\text{pred}}$ and $\text{cvc4}_{\text{term}}$ are obtained from the CVC4 group~\cite{termEQ}. This benchmark also contains some string-number conversion constraints.
\end{itemize}

In the second experiment, we compare with tools supporting to string-number conversion on the  benchmarks collected from the symbolic executor \texttt{Py-Conbyte}\footnote{\url{https://github.com/spencerwuwu/py-conbyte}}, which has the supports of string-number conversion. We run it on several examples collected from the LeetCode platform and from Python core libraries.

@@

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. The number of failed cases is only a half to Z3, which is ranked the 2nd. In fact, most of the failed tests comes from the analysis of {lib-ipaddress\_ip\_int\_from\_string}, which converts an IPv6 address to a number. This function has a very long execution path and hence the size of each SMT file from this analysis is roughly 1MB. If we exclude results on this benchmark, then {\tool} has only $185$ failed case. This is significantly better than the 2nd place tool Z3, which fails $893$ cases.

>>>>

\begin{table}[h]
\centering
\caption{Results of Z3-Trau, CVC4, and Z3 on Basic String Constraint benchmarks (numbers with * will be updated later)}
\scalebox{0.7}{
\begin{tabular}{|l r | r r r r r r r|}
\hline
\multicolumn{2}{|c}{}                   & {\tool} & CVC4  &Z3 & Z3-str3 & Trau+ & ABC & Ostrich \\ \hline
\multirow{3}{*}{PyEx}		& sat      & 19586*   & 19763* & 18359 &   3024*  & 19149 & 0 & 111 \\ 
							& unsat    &  3858*   &  3834* &  3851 &   3839*  & 3828 & 0 & 871 \\
							& $\times$ &  1977*   &  1824* &  3211 &  18558*  & 2444 & 25421 & 24439 \\ \hline
\multirow{3}{*}{APLAS}		& sat      &   128*   &   205* &    19 &     38*  & 132 & 289 & 0 \\
							& unsat    &   287*   &   211* &   100 &    111*  & 82 & 2 & 1 \\
							& $\times$ &   185*   &   174* &   481 &    451*  & 386 & 309 & 599 \\ \hline
\multirow{3}{*}{LeetCode}	& sat      &   856*   &   860* &   881 &    670*  & 778 & 0 & 158 \\
							& unsat    &  1784*   &  1785* &  1785 &   1780*  & 1827 & 0 & 1618 \\
							& $\times$ &    26*   &    21* &     0 &    216*  & 61 & 2666 & 890 \\ \hline
\multirow{3}{*}{StringFuzz}	& sat      &   502*   &   677* &   311 &    493*  & 688 & 439 & 0 \\
							& unsat    &   294*   &   240* &   205 &    190*  & 339 & 158 & 0 \\
							& $\times$ &   269*   &   148* &   549 &    382*  & 38 & 468 & 1065 \\\hline
\multirow{3}{*}{cvc4\textsubscript{pred}} & sat & 16* & 11* &   12 &      8*  & 247 & 316 & 21 \\
							& unsat    &   814*   &   818* &   808 &    772*  & 500 & 443 & 17 \\
							& $\times$ &     5*   &     6* &    15 &     55*  & 88 & 76 & 797 \\ \hline
\multirow{3}{*}{cvc4\textsubscript{term}} & sat & 13* & 8* &     6 &     16*  & 311 & 576 & 45 \\
							& unsat    &  1030*   &   936* &  1022 &    958*  & 596 & 349 & 22 \\
							& $\times$ &     2*   &   101* &    17 &     71*  & 137 & 120 & 978 \\ \hline
\multirow{3}{*}{SLOG}  		& sat      & 0* & 1309* & 0 & 0* & 1228 & 1036 & 1299 \\
							& unsat    & 3391* & 2082* & 0 & 2205* & 2079 & 1972 & 2079 \\
							& $\times$ & 0* & 0* & 3391 & 1186* & 84 & 383 & 13 \\ \hline
							\hline
\multirow{3}{*}{Total}  		& sat      & 0* & 1309* & 0 & 0* & 1228 & 1036 & 1299 \\
& unsat    & 3391* & 2082* & 0 & 2205* & 2079 & 1972 & 2079 \\
& $\times$ & 0* & 0* & 3391 & 1186* & 84 & 383 & 13 \\\hline
\end{tabular}}
\label{table:base_benchmark}
\end{table}

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, and Z3 on Basic String Constraints + String-Number Conversion benchmark (numbers with * will be updated later)}
\scalebox{0.8}{
\begin{tabular}{|l r | r r r r|}
\hline
\multicolumn{2}{|c}{}                   & {\tool} & CVC4   &    Z3  & Z3-str3 \\ \hline
\multirow{3}{*}{String-Number-Leetcode}		& sat      &   3294*  &  2185* &  2731* &    422* \\ 
							& unsat    &  17088*  & 16377* & 16832* &  16034* \\
							& $\times$ &   1191*  &  3011* &  2010* &   5117* \\ \hline 
\multirow{3}{*}{String-Number-PythonLib}		& sat      &   3294*  &  2185* &  2731* &    422* \\ 
& unsat    &  17088*  & 16377* & 16832* &  16034* \\
& $\times$ &   1191*  &  3011* &  2010* &   5117* \\ \hline
\multirow{3}{*}{String-Number-JavaScript}		& sat      &   3294*  &  2185* &  2731* &    422* \\ 
& unsat    &  17088*  & 16377* & 16832* &  16034* \\
& $\times$ &   1191*  &  3011* &  2010* &   5117* \\ \hline 							
							\hline
\multirow{3}{*}{Total}	& sat & 622*  &  330* &  100* &  85* \\ 
& unsat    &  1054*  & 780* & 1043* &  634* \\
& $\times$ &  2*  &  568* &  535* & 959* \\ \hline
\end{tabular}}
\label{table:str_int_benchmark}
\end{table}

\hide{%3rd version
\begin{table}[t]
	\centering
	\caption{Results of {\tool}, cvc4, and z3 on str\_int benchmark (numbers with * will be updated later)}
	\scalebox{0.7}{
		\begin{tabular}{l r | r r r r}
			\hline
			\multicolumn{2}{c}{}                   & {\tool} & CVC4   &    Z3  & Z3-str3 \\ \hline
			\multirow{3}{*}{String-Number}		& sat      &   3294*  &  2185* &  2731* &    422* \\ 
			& unsat    &  17088*  & 16377* & 16832* &  16034* \\
			& $\times$ &   1191*  &  3011* &  2010* &   5117* \\ \hline \hline

				\multirow{3}{*}{leetcode-addStrings}	& sat & 622*  &  330* &  100* &  85* \\ 
				& unsat    &  1054*  & 780* & 1043* &  634* \\
				& $\times$ &  2*  &  568* &  535* & 959* \\ \hline
				\multirow{3}{*}{leetcode-add\_Binary}	& sat & 520*  &  530* &  531* &  11* \\ 
				& unsat    &  1351*  & 1091* & 1091* &  1094* \\
				& $\times$ &  11*  &  261* &  260* & 777* \\ \hline
				\multirow{3}{*}{leetcode-numDecodings}	& sat & 166*  &  107* &  151* &  50* \\ 
				& unsat    &  273*  & 269* & 283* &  269* \\
				& $\times$ &  92*  &  155* &  97* & 212* \\ \hline
				\multirow{3}{*}{leetcode-restoreIpAddresses}	& sat & 811*  &  462* &  864* &  55* \\ 
				& unsat    &  13648*  & 13540* & 13648* &  13342* \\
				& $\times$ &  54*  &  511* &  1* & 1116* \\ \hline
				\multirow{3}{*}{leetcode-validIPAddress}	& sat & 54*  &  57* &  58* &  8* \\ 
				& unsat    &  19*  & 19* & 19* &  19* \\
				& $\times$ &  4*  &  1* &  0* & 50* \\ \hline
				\multirow{3}{*}{leetcode-validWordAbbreviation}	& sat & 183*  &  193* &  188* &  8* \\ 
				& unsat    &  24*  & 11* & 21* &  16* \\
				& $\times$ &  39*  &  42* &  37* & 222* \\ \hline
				\multirow{3}{*}{lib-datetime\_parse\_hh\_mm\_ss\_ff}	& sat &  133*  & 133* &  133* &  88* \\ 
				& unsat    &  41*  & 41* & 41* &  43* \\
				& $\times$ &  0*  &  0* &  0* & 43* \\ \hline
				\multirow{3}{*}{lib-datetime\_parse\_isoformat\_date}	& sat & 32*  &  32* &  32* &  23* \\ 
				& unsat    &  0*  & 0* & 0* &  0* \\
				& $\times$ &  0*  &  0* &  0* & 9* \\ \hline
				\multirow{3}{*}{lib-distutils\_get\_build\_version}	& sat & 19*  &  19* &  19* &  4* \\ 
				& unsat    &  24*  & 24* & 24* &  24* \\
				& $\times$ &  0*  &  0* &  0* & 15* \\ \hline
				\multirow{3}{*}{lib-email\_parsedate\_tz}	& sat & 68*  &  72* &  72* &  30* \\ 
				& unsat    &  138*  & 138* & 138* &  138* \\
				& $\times$ &  4*  &  0* &  0* & 42* \\ \hline
				\multirow{3}{*}{lib-http\_parse\_request}	& sat & 24*  &  24* &  24* &  8* \\ 
				& unsat    &  9*  & 9* & 9* &  7* \\
				& $\times$ &  0*  &  0* &  0* & 18* \\ \hline
				\multirow{3}{*}{lib-ipaddress\_ip\_int\_from\_string}	& sat & 595*  &  204* &  480* &  18* \\ 
				& unsat    &  427*  & 373* & 431* &  357* \\
				& $\times$ &  1006*  &  1451* &  1117* & 1653* \\ \hline
				\multirow{3}{*}{lib-nntplib\_parse\_datetime}	& sat & 4*  &  8* &  0* &  0* \\ 
				& unsat    &  0*  & 0* & 0* &  0* \\
				& $\times$ &  4*  &  0* &  8* & 8* \\ \hline
				\multirow{3}{*}{lib-smtpd\_parseargs}	& sat & 27*  &  27* &  27* &  24* \\ 
				& unsat    &  72*  & 72* & 72* &  66* \\
				& $\times$ &  0*  &  0* &  0* & 9* \\ \hline
				\multirow{3}{*}{lib-wsgiref\_check\_status}	& sat & 10*  &  10* &  10* &  6* \\ 
				& unsat    &  9*  & 9* & 9* &  9* \\
				& $\times$ &  0*  &  0* &  0* & 4* \\ \hline
				% \multirow{4}{*}{filtered\_str\_int}	& sat      &   3281*  &  2210* &  2729* &    424* \\
				% 									& unsat    &   2912*  &  2211* &  2655* &   1944* \\
				% 									& $\times$ &   1203*  &  2975* &  2012* &   5028* \\ \hline
			}
		\end{tabular}
		\label{table:str_int_benchmark}
	\end{table}
}

@@

We have implemented our string constraint solving procedure, in a tool called {\tool}, on top of the SMT solver Z3~\cite{z3} as a theory solver. In this way we can concentrate on solving constraints connected by conjunctions and let Z3 to handle other boolean connectives. It also enable us to solve not only formulas over string constraints, but also the combination of string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulas.

@@

We perform two sets of experiments. In the first experiment, we compare with other tools on benchmarks over basic string constraints. We use the following benchmark examples in this experiments:
\begin{itemize}
	\item PyEx~\cite{pyex} is from running the symbolic executor PyEx over some Python packages.
	\item APLAS~\cite{aplas} includes 600 hand-drafted tests consisting of only equality and integer length constraints.
	\item LeetCode is from running PyEx over some sample code collected from the LeetCode~\cite{LeetCode} website.
	\item StringFuzz~\cite{StringFuzz}  is generated from the fuzz testing tool of the same name.
	\item $\text{cvc4}_{\text{pred}}$ and $\text{cvc4}_{\text{term}}$ are obtained from the CVC4 group~\cite{termEQ}. These benchmarks contain a small amount of string-number conversion constraints (< $5\%$).
\end{itemize}

In the second experiment, we compare with tools supporting to string-number conversion on the  benchmarks collected from the symbolic executor \texttt{Py-Conbyte}\footnote{\url{https://github.com/spencerwuwu/py-conbyte}}, which has the supports of string-number conversion. We run it on several examples collected from the LeetCode platform and from Python core libraries. We also have examples that encode execution paths of some JavaScript programs (the Luhn algorithm and some array manipulation).

@@

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. The total number of failed cases is only a half to Z3, which is ranked the 2nd. In fact, most of the failed tests comes from the analysis of one Python core library function that converts an IPv6 address to a number. We generated 2028 tests from this function, and among them around 1000 cases that are too difficult for all solvers.  If we exclude results on this function, then {\tool} has only $185$ failed case in total. This is significantly better than the 2nd place tool Z3, which fails $893$ cases in total.
